A Data-Driven Framework for Neural Field Modelling

### 2010-11-16 ltx/first_order_framework/LaTeX_Manuscript.tex
minor correction
changed figure 13
generalise the model for non isotropic case

where $\boldsymbol{\psi}(\mathbf{r},\mathbf{r}')$ is a vector of Gaussian basis functions and $\boldsymbol{\theta}$ is a vector of scaling parameters. By assuming an isotropic connectivity structure, the kernel basis functions can be written as $\boldsymbol{\psi}(\mathbf{r}-\mathbf{r}')$. We will assume that we know the parametric form of the connectivity basis functions, where the scaling parameters $\boldsymbol{\theta}$ are unknown. Each connectivity basis function can, individually, be considered a layer in the Wilson and Cowan model. Typically, the Mexican hat connectivity kernel is comprised of two basis functions representing local excitation and surround inhibition. To demonstrate the flexibility of the estimation framework, a third basis function is incorporated, corresponding to weaker mid-range excitation. Rearranging the order of the spatial convolution in equation~\ref{DiscreteTimeModel} and substituting in equations~\ref{DefFieldDecomp} and~\ref{DefKernelDecomp} we get 

@@

The analytic derivation of the inner product of \emph{n}-dimensional Gaussians, which was used for calculating $\mathbf\Gamma$ is provided in the supplementary material. Equation~\ref{eq:ReducedForm} can be simplified by exploiting the isotropy of the kernel and, hence, the basis functions used to represent it where

@@

	\boldsymbol{\psi} (\mathbf{r}-\mathbf{r}') = \boldsymbol{\psi} (\mathbf{r}'-\mathbf{r}).

@@

To make the simplification, we define

@@

	\boldsymbol{\Psi}(\mathbf{r}') \triangleq T_s\boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top} (\mathbf{r}'-\mathbf{r}) \, d\mathbf{r}},

@@

which is a constant $L \times n_{\theta}$ matrix that is defined analytically (see the supplementary material for analytic convolution of two Gaussians), where $L$ is the number of basis functions (states) and $n_{\theta}$ is the number of connectivity kernel basis functions. Substituting equation~\ref{eq:DefPsi} into equation~\ref{eq:ReducedForm} gives
	\mathbf{x}_{t+1} &=& \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \, d\mathbf{r}' \boldsymbol{\theta} \nonumber \\ &+& \xi\mathbf{x}_t 
+ \boldsymbol{\Gamma}^{-1} \int_\Omega{\boldsymbol{\phi}(\mathbf{r})e_t(\mathbf{r}) \, d\mathbf{r}}.
This substitution exploits the isotropy of the connectivity kernel basis functions, allowing us to swap the convolution in equation~\ref{reduced continuous model}, which would be computationally demanding, with the convolution in equation~\ref{eq:DefPsi}, which can be calculated analytically. This is of great importance as it lowers the computational demands of the estimation algorithm and provides a dramatic increase in estimation speed.

@@

% The size of the squares in the figure show the magnitude of the log of the error. The squares are black if the log of the error is negative (the RMSE is $<1$) and the squares are white if the log of the error is positive (the RMSE is $>1$). The largest black squares indicate the best performance and the largest white squares indicate the worst performance for the parameter combinations (\dean{need some thing more quantitative}). 
From Figure~\ref{fig:Figure12} it can be seen that the estimation accuracy depends the parameters of the sigmoid function. \dean{This next bit is CRAP (please help):} When the threshold is negative or the slope is very gentle, all of the activity is being passed through the sigmoid function. However, as the slope is increased or the threshold is increased, more of the field is within the saturating region of the activation function. This is illustrated in Figure~\ref{fig:sigmoidfunctions}, where the sigmoids from the bottom left corner through to near top right corner of Figure~\ref{fig:sigmoidfunctions} are plotted. This appears to have a detrimental effect on the estimation accuracy.

@@

\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/RMSESensor.pdf}
\end{center}
\caption{{\bf RMSE of the kernel estimates}. RMSE of the kernel for different values of sensor width, $\sigma_m$, distance between adjacent sensors is $\Delta_y=1.5$. Mean kernel estimates are obtained over 50 realisations in each case.}
\label{fig:Figure13}
\end{figure*}
\section*{Tables}

>>>>

where $\boldsymbol{\psi}(\mathbf{r},\mathbf{r}')$ is a vector of Gaussian basis functions and $\boldsymbol{\theta}$ is a vector of scaling parameters. \parham{By exploiting the isotropy of the connectivity basis functions, the kernel basis functions can be written as  
$\boldsymbol{\psi}(\mathbf{r}-\mathbf{r}')$.} By assuming an isotropic connectivity structure, the kernel basis functions can be written as $\boldsymbol{\psi}(\mathbf{r}-\mathbf{r}')$. We will assume that we know the parametric form of the connectivity basis functions, where the scaling parameters $\boldsymbol{\theta}$ are unknown. Each connectivity basis function can, individually, be considered a layer in the Wilson and Cowan model. Typically, the Mexican hat connectivity kernel is comprised of two basis functions representing local excitation and surround inhibition. To demonstrate the flexibility of the estimation framework, a third basis function is incorporated, corresponding to weaker mid-range excitation. Rearranging the order of the spatial convolution in equation~\ref{DiscreteTimeModel} and substituting in equations~\ref{DefFieldDecomp} and~\ref{DefKernelDecomp} we get 

@@

The analytic derivation of the inner product of \emph{n}-dimensional Gaussians, which was used for calculating $\mathbf\Gamma$ is provided in the supplementary material. Equation~\ref{eq:ReducedForm} \parham{can be simplified by exploiting the isotropy of the kernel  basis functions used to represent the kernel , and hence,

@@

	\psi_i (\mathbf{r}-\mathbf{r}') = \psi_i (2c_i+\mathbf{r}'-\mathbf{r}).

@@

where $c_i$ is the center of the $i^{th}$ basis function of the kernel decomposition. To make the simplification, we define

@@

	\left[ \boldsymbol\Psi(\mathbf{r}')\right]_{:i}  \triangleq T_s\boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\psi_i (2c_i+\mathbf{r}'-\mathbf{r})\textrm{d}\mathbf{r}},

@@

where $\left[ \boldsymbol\Psi(\mathbf{r}')\right]_{:i}$ denotes the $i^{th}$column of $\boldsymbol{\Psi}(\mathbf{r}')$ which is a constant $L \times n_{\theta}$ matrix that is defined analytically (see the supplementary material for analytic convolution of two Gaussians), where $L$ is the number of basis functions (and states) and $n_{\theta}$ is the number of connectivity kernel basis functions.}

can be simplified by exploiting the isotropy of the kernel and, hence, the basis functions used to represent it where

\begin{equation}
	\boldsymbol{\psi} (\mathbf{r}-\mathbf{r}') = \boldsymbol{\psi} (\mathbf{r}'-\mathbf{r}).
\end{equation}
To make the simplification, we define
\begin{equation}\label{eq:DefPsi}
	\boldsymbol{\Psi}(\mathbf{r}') \triangleq T_s\boldsymbol{\Gamma}^{-1}\int_\Omega {\boldsymbol{\phi}(\mathbf{r})\boldsymbol{\psi}^{\top} (\mathbf{r}'-\mathbf{r}) \, d\mathbf{r}},
\end{equation}
which is a constant $L \times n_{\theta}$ matrix that is defined analytically (see the supplementary material for analytic convolution of two Gaussians), where $L$ is the number of basis functions (states) and $n_{\theta}$ is the number of connectivity kernel basis functions. Substituting equation~\ref{eq:DefPsi} into equation~\ref{eq:ReducedForm} gives
\begin{eqnarray}
	\mathbf{x}_{t+1} &=& \int_\Omega \boldsymbol{\Psi}(\mathbf{r}') f(\boldsymbol{\phi}^{\top}(\mathbf{r}')\mathbf{x}_t) \, d\mathbf{r}' \boldsymbol{\theta} \nonumber \\ &+& \xi\mathbf{x}_t 
+ \boldsymbol{\Gamma}^{-1} \int_\Omega{\boldsymbol{\phi}(\mathbf{r})e_t(\mathbf{r}) \, d\mathbf{r}}.
\end{eqnarray}
This substitution exploits the isotropy of the connectivity kernel basis functions, allowing us to swap the convolution in equation~\ref{reduced continuous model}, which would be computationally demanding, with the convolution in equation~\ref{eq:DefPsi}, which can be calculated analytically. This is of great importance as it lowers the computational demands of the estimation algorithm and provides a dramatic increase in estimation speed.

@@

% The size of the squares in the figure show the magnitude of the log of the error. The squares are black if the log of the error is negative (the RMSE is $<1$) and the squares are white if the log of the error is positive (the RMSE is $>1$). The largest black squares indicate the best performance and the largest white squares indicate the worst performance for the parameter combinations (\dean{need some thing more quantitative}). 
From Figure~\ref{fig:Figure12} it can be seen that the estimation accuracy depends on the parameters of the sigmoid function. \dean{This next bit is CRAP (please help):} When the threshold is negative or the slope is very gentle, all of the activity is being passed through the sigmoid function. However, as the slope is increased or the threshold is increased, more of the field is within the saturating region of the activation function. This is illustrated in Figure~\ref{fig:sigmoidfunctions}, where the sigmoids from the bottom left corner through to near top right corner of Figure~\ref{fig:sigmoidfunctions} are plotted. This appears to have a detrimental effect on the estimation accuracy.

@@

\begin{figure*}[!ht]
\begin{center}
\includegraphics{./Graph/pdf/Figure13.pdf}
\end{center}
\caption{{\bf RMSE of the kernel estimates}. RMSE of the kernel for different values of sensor width, $\sigma_m$, distance between adjacent sensors is $\Delta_y=1.5$. Mean kernel estimates are obtained over 50 realisations in each case.}
\label{fig:Figure13}
\end{figure*}
\section*{Tables}
